%%
%% This is a skeleton file demonstrating the use of the curtinThesis.cls class file
%% (requires curtinThesis.cls and LaTeX2e format). And can be used as infrastructure
%% to build your thesis.
%%
%% --------------------------------------------------------------------------------
%%
%%   This work is licensed under the conditions of the Creative
%%   Commons Attribution 4.0 International (CC BY 4.0) license:
%%
%%		<https://creativecommons.org/licenses/by/4.0/>
%% 
%%   This work may be remixed, tweaked, and build upon, even
%%	 commercially, as long as the original Author is credited
%%   for the creation of the original work. 
%% 
%%   This code is offered as-is without any warranty either expressed
%%   or implied; User assumes all risk.
%%
%% --------------------------------------------------------------------------------
%% 
%%  Usage: (this usage block can be removed after reading)
%%  ------
%%  Define you thesis title, name, [optional] auxiliary association,
%%	graduation date, [optional] quote, [optional] linkcolor,
%%  and [optional] cite color under the ``Define Globals Here''
%%  section.
%%
%%	Adjust, edit, or extend the Chapter files in the chapters folder
%%	to match your thesis content accordingly. If one decides to alter
%%	file names, change the names in this template file as well. 
%%
%%	One can select APA, Vancouver, Chicago, or Harvard referencing in the class
%%	options using apa, vancouver, chicago, or harvard respectively. For other class
%%	options available, please refer to the curtinThesis class or the book class.
%%
%%	All functionality in this template is commented, the minimal action
%%  required is to follow the above steps and to build the document. 
%%
%%  Alterations to the template file, other than the global definitions,
%%	is not required, but can be done to suit your needs and/or requirements.
%%
%%  When changing reference styles, please make sure to clean the auxiliary
%%	files to prevent any package clashes.
%%
%% --------------------------------------------------------------------------------
%%

\documentclass[apa,colorlinks,emptypage]{curtinThesis}
\usepackage[all]{unitabbrev} %optional package that includes standard unit abbreviations
\usepackage{adsshort} %optional package that includes abbreviations endorsed by the ADS
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{amsmath,mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}

%\usepackage[ruled,vlined]{algorithm2e}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Define Globals Here  %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%

% define your thesis title, name, [optional: auxiliary association], and graduation date
\thesistitle{Bayesian Hierarchical Modelling of Equipment Reliability in Mining: A Pragmatic Approach} %your thesis title

\authorname{Ryan K. Leadbetter} % your name e.g. John A. Doe


\auxassoc{Centre for Transforming Maintenance Through Data Science} % remove, comment, or leave empty if no auxiliary association

\graduationdate{November 2023} % your expected graduation/hand-in date

% [optional: define a quote to appear in thesis - \thesisquote{"the quote"}{"the author"}]
\thesisquote{The Quote}{AUTHOR} % remove, comment, or leave empty when not used

% define your link and citecolor - standard is cyan
\linkcolor{red} % remove, comment, or leave empty if not specifying different color
\citecolor{black} % remove, comment, or leave empty if not specifying different color

% set the metadata of the .pdf file correctly
\setmetadata % make sure the metadata of your pdf file is according to your work.

\begin{document}

\createtitlepage % generate a title page

\plagiarismstatement % generate a plagiarism statement

\showthesisquote % if selected, show thesis quote

% include the acknowledgements
\include{chapters/acknowledgements}

% include the abstract
\include{chapters/abstract}

% include the table of contents
\includetoc % neatly adds the table of contents

% include the list of figures
\includelof % neatly adds the list of figures


\arabicnumbering{1} % set pagenumbers to arabic and start at 1 - change number in brackets to change start number

% include chapters - adjust the chapter files in chapters folder - remove or extend where necessary
\include{chapters/chapter1}

\part{Part one: lifetime analysis}\label{part:one}

A preamble about which chapters have been published and which chapters came from industry placements.

``Since the initial approach was published in Leadbetter et al. (2021), I have taken an alternative approach that does\ldots''

\include{chapters/chapter2}
\include{chapters/chapter3}

\part{Part two: Degradation modelling}\label{part:two}

In this second part of the thesis, I focus on degradation modelling---primarily using the gamma stochastic process. The more foundational work in this second half of the thesis---chapters~\ref{chap:chapter4} and~\ref{chap:chapter5}---arose out of necessity to apply the gamma process to industry data from the Centre's industry partners. Particularly, the work was motivated by two applied problems: modelling the degradation of many pipe spools in a bauxite processing plant and modelling the wear of an overland iron ore conveyor's belt---the latter was in one of my industry placements. The pipe degradation project, undertaken by  Gabriel Gonz\'{a}lez Caceres and Aloke Phatak, looked at the degradation in wall thickness of pipe spools. The gamma process should theoretically be a good model for the erosion of the inside of the spools. However, the degradation signals for each spool are very noisy and consist of only a few measurements; therefore, there was a need to accommodate noise in the gamma process model and to borrow information between similar units in order to estimate the parameters of the model. The belt wear project was an industry placement where I was asked to analyse the belt wear of all conveyors at an industry partner mine, processing facility, and port and devise a better method for modelling belt wear. To analyse the many wear profiles, I used functional data analysis techniques. This led to taking the same approach when modelling the belts' degradation. The gamma process should also be a good model for the degradation of the belt since the degradation mechanism is wear. However, similar to the pipe spool problem, the signals of the spline coefficients were noisy and sparsely observed.

In the first chapter of this part, chapter~\ref{chap:chapter4}, I demonstrate how the gamma process can be extended to incorporate noisy degradation measurements through Bayesian hierarchical modelling. Then, in chapter~\ref{chap:chapter5}, I show how unit-to-unit variability (which facilitates the borrowing of information between units) can be included in the proposed noisy gamma process model. The work in these two chapters has been submitted for publication as the manuscript \textit{``Bayesian hierarchical modelling of noisy gamma processes: model formulation, identifiability, model fitting, and extensions to unit-to-unit variability''} to Applied Models in Business and Industry. The chapters present an expansion of the manuscript by tying the supplementary material submitted alongside the manuscript into the main analysis and discussion and providing a deeper discussion of the two main topics. There is also a larger focus on the computational details of the analysis.

In chapter~\ref{chap:chapter6}, I present the belt wear model developed as part of the belt wear industry placement, which utilises the noisy gamma process model. The work in chapter~\ref{chap:chapter6} has been submitted for publication to the Annals of Applied Statistics as the manuscript \textit{``Functional degradation modelling of the wearing surface of conveyor belting using Bayesian hierarchical modelling and Gamma processes''}. In addition to the work presented in the manuscript, chapter~\ref{chap:chapter6} provides a side-by-side comparison of the gamma process-based belt wear model from the manuscript with a linear general path-based model.

The e-prints for the two papers have been uploaded to arXiv. The full references for the manuscripts are
\begin{enumerate}
  \item Leadbetter, R., Gonz\'{a}lez C\'{a}ceres, G., \& Phatak, A. (2024). \textit{Bayesian hierarchical modelling of noisy gamma processes: Model formulation, identifiability, model fitting, and extensions to unit-to-unit variability.}
  \item Leadbetter, R. \& Phatak, A. (2024). \textit{Functional degradation modelling of the wearing surface of conveyor belting using Bayesian hierarchical modelling and Gamma processes}
\end{enumerate}

\include{chapters/chapter4}
\include{chapters/chapter5}
\include{chapters/chapter6}

\include{chapters/chapter7}

% start your appendices
\begin{customappendix}{} % add \alphanumbering between brackets if you want to have alphabetic numbered appendices
\include{chapters/appendixA}
\include{chapters/copyright}

\end{customappendix}

\backmatter{}

\bibliography{references} % change 'references' to your .bib file
\ackstatement % statement to acknowledge the effort put into acknowledging the authors
\end{document}
