%%
%% This is a file demonstrating the use of the acknowledgement file in the Curtin thesis skeleton
%% file. And can be used as infrastructure to build your thesis.
%%

\chapter*{Acknowledgements} \addcontentsline{toc}{chapter}{Acknowledgements}

First and foremost, to my primary supervisor, Aloke Phatak. Thank you for your guidance and mentorship. It has been an honour. Being witness to your rigour, passion, kindness, and thirst for knowledge has changed me fundamentally.

I would also like to thank my co-supervisors, Michael Small and Chris Aldrich (and, for a short time, Darfiana Nur).

My PhD was supported by the Center for Transforming Maintenance through Data Science (CTMTDS), for which I am extremely grateful. Thank you to all of the passionate and brilliant people who make CTMTDS what it is: a community. And to the person who made it that most, Natasha Barlett, I have never seen someone go so above and beyond their job description. Thank you for being a mentor and friend, for always having an open door for me or anyone else (not that you had much of a choice in an open-plan office), and for always checking in.

To my 216 office mates Sandy, Gabriel, Yingying, Ponpot, and Srimali, thank you for your fellowship: the lunch chats; the coffee rants; the travel adventures; the pub celebrations. I have loved watching your journeys in parallel to my own, I cannot wait to see where your careers go.

To my two `unofficial' UWA supervisors, Melinda and Adriano, thank you for your mentorship, the opportunities that you have given me, and the feedback on my research.

To the small Bayesian community at Curtin: Arash, Suman, Zanglong, and Shih Ching. It was a pleasure to learn from and alongside you, and thank you for volunteering your time for me to bounce ideas off you. 

To my industry supervisors and collaborators: Chris, Julian, Michael, and Mittin. Thank you for seeing the value in academic research and investing your most valuable currency in it: your time.

To Saf and Hutch, thank you for your enthusiasm to contribute to the practical side of this thesis. Hutch, getting to go up to site and get hands-on with the conveyors with you was invaluable in the applied component of the work contained in these pages.

To my family, Anne, Russ, and Ellie (and Charlie). This thesis would not only not exist without your love and support, but I would not have even attempted it. Mum, my other ghost supervisor, thank you for helping me navigate this journey. With your passion for science and research, I think it was inevitable that I would end up here.

Lastly, to Jody, my better half. I cannot thank you enough for the support that you have so unconditionally given, especially over these last few months. Thank you for always making sure I enjoy the finer things life has to offer outside the world of work and academia.