
\chapter*{Thesis Publications} \addcontentsline{toc}{chapter}{Thesis Publications} \label{chap:publications}

A portion of the works in this thesis have been published or is in the process of being published. The groundwork for Chaps.~\ref{chap:chapter2} and~\ref{chap:chapter3} on heavily censored lifetime data were published as \citet{leadbetter2021}[1]. However, Chaps.~\ref{chap:chapter2} and~\ref{chap:chapter3} present an alternative approach to the original manuscript. The work in Chapts.~\ref{chap:chapter4} and~\ref{chap:chapter5} focussing on the noisy gamma process have been published as \citet{leadbetter2024}[2] and the work in Chap.~\ref{chap:chapter6} on conveyor belt wear will be published as \citet{leadbetter2025}[3].

\begin{enumerate}
  \item \textbf{Leadbetter, R.}, Phatak, A., Polpo, A. \& Hodkiewicz, M. (2021) \textit{Informative Bayesian Survival Analysis to Handle Heavy Censoring in Lifetime Data}, International Conference on Maintenance and Intelligent Asset Management (ICMIAM), Ballarat, Australia, pp. 1-6,\\ doi: 10.1109/ICMIAM54662.2021.9715184.
  \item \textbf{Leadbetter, R.}, Gonz\'{a}lez C\'{a}ceres, G., \& Phatak, A. (2024). \textit{Bayesian hierarchical modelling of noisy gamma processes: Model formulation, identifiability, model fitting, and extensions to unit-to-unit variability.}, arXiv. https://arxiv.org/abs/2406.11216 \\
  (Submitted to and under review by \textit{Applied Stochastic Models in Business and Industry}.)
  \item \textbf{Leadbetter, R.} \& Phatak, A. (2024). \textit{Functional degradation modelling of the wearing surface of conveyor belting using Bayesian hierarchical modelling and Gamma processes.} \\
  (Manuscript in progress.)
\end{enumerate}